SatWins: A Sat-based tool to Optimal Instantiations of Abstract Workflows

### 2010-01-25 publicaciones/demo_eswc10/paperdemo.tex
demo_eswc10

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{url}

@@

\begin{document}

\allowdisplaybreaks
\title{Optimal Instantiation of Abstract Workflows using Logic and Circuits}
\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
\institute{Departamento de Computaci\'on \\
           Universidad Sim\'on Bol\'{\i}var \\
           Caracas 89000, Venezuela \\
           \url{{idaniel,bonet,mvidal}@ldc.usb.ve}}
\maketitle

\begin{abstract}
The typical Service Oriented Architecture consists of two main layers.
The concrete layer with concrete services whose functionalities are
described in terms of pre- and post-conditions and non-functional
properties in terms of QoS parameters, and the abstract layer with
software applications whose functionalities are described in terms
of abstract workflows and non-functional properties in terms of
restrictions on QoS.
During the execution of a workflow, the abstract services are instantiated
into concrete services that meet the functional and non-functional
requirements. This instantiation, that must be done on-the-flight,
consists in a search on a combinatorial space of possibilities.
In this paper, we propose a framework to efficiently solve the
instantiation problem that is firmly grounded on logic.
The framework adopts the Local-As-View approach in which
the functionality of concrete services is described using views
of abstract services, the quality of an instantiation as an
overall utility function that combines the different QoS parameters,
and the abstract workflows as conjunctive queries on abstract services.
Using this representation, the problem of workflow instantiation 
is cast as a problem of query rewriting, from the area of integration
systems.
Then, building on related work, we devise an encoding of the
workflow instantiation problem as a logical theory whose models
are in correspondence with the instantiations of the workflow,
and the best ranked models are in correspondence with the
optimal instantiations of the workflow.
Thus, by exploiting known properties of logical theories in 
d-DNNF format, we provide an efficient and scalable solution
to the workflow instantiation problem.
The approach does not only scale up to large instances, as the
experimental results show, but is also sound and complete since,
being founded on logic, is amenable to formal analysis.
\end{abstract}

\section{Introduction}

Under the umbrella of the Semantic Web, and supported by Service Oriented
Architectures (SOA), the number of Web data sources and services has exploded
during the last few years. For example, the molecular biology database
collection currently contains 1,170 databases~\cite{Galperin09}, a number
that is 95 more than the previous year~\cite{Galperin2008} and 110 more
than two years ago~\cite{Galperin2007}; tools and services, as well as the
number of instances published by these resources, follow a similar
progression~\cite{Benson07}. Thanks to this wealth of resources, users'
tendency is to rely more on automatic methods for handling them such as
data retrieval from public sources and analysis using Web
tools or services composed in complex workflows.

The typical SOA consists of two layers. The concrete layers that is
made of concrete services whose functionalities are described in
terms of pre- and post-conditions and non-functional properties in
terms of QoS parameters, and the abstract layer that is made of 
software applications whose functionalities are described in terms
of abstract workflows and non-functional properties in terms of
restrictions on QoS.
The execution of an abstract workflow involves the instantiation
of the abstract services into concrete services that meet the
functional and non-functional requirements. This instantiation
process can be seen as a search for a target instantiation on
the combinatorial space of all valid instantiations.
Thus, one is interested in efficient techniques for performing
this search that are able to scale up as the number of concrete
services or the complexity of the workflow increases.
We call the problem of instantiating a given abstract workflow
with concrete services, from a given pool of concrete services,
such that certain QoS demands are meet, the Workflow Instantiation
Problem (WIP).

In this paper, we consider a restricted version of WIP that adopts
the Local-As-View (LAV) approach \cite{levy:bucket}.
In LAV, all elements in a problem are specified with a common language
that is grounded on \emph{abstract services} such that concrete
services are described as views of abstract services, the quality
of an instantiation as an overall utility function that combine the
different QoS parameters, and the abstract workflow as a conjunctive
query on the abstract services; this representation is similar to the
one that is semi-automatically generated for the DEIMOS system
\cite{AmbiteISWC09}.
In this version of WIP, the rules that define workflows (conjunctive queries)
and concrete services (views) are created in a way that all the 
functional restrictions on pre- and post-conditions of services
and their combinations are satisfied, and the QoS measures are
represented by annotating each concrete service description with
a real number that represent the overall QoS utility of the service.

Under these assumptions, the WIP can be cast as the well-known Query
Rewriting Problem (QRP) for LAV which is central to integration systems
\cite{halevy:survey}. The QRP consists of a conjunctive query that must be
answered in terms of views in which the query and the views are described
using LAV with abstract relations.
This problem is important in the context of data integration
\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance
\cite{AfratiLU07,levy:bucket}, and several approaches have been defined
that scale to a large number of views
\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.

The recent approach of Arvelo et al.\ \cite{arvelo:aaai06} is based 
on the efficient enumeration of models of a propositional logic theory. 
Given a QRP, a logical theory is constructed such that each model of
the theory encodes a valid rewriting, and thus all valid rewritings
are obtained by enumerating all models of the theory. This enumeration
can be efficiently performed if the logical theory is in certain 
normal form called deterministic and decomposable negation normal
form (d-DNNF) \cite{darwiche:d-dnnfs}. Thus, the approach consists in
transforming (called compiling in the field of knowledge compilation)
the logical theory into d-DNNF format for enumerating its models efficiently.

Yet d-DNNF theories not only support the efficient enumeration of
models but other operations too. Among them, the enumeration of
the best ranked models can also be performed efficiently on d-DNNF.
Given a literal ranking function $r(\ell)$ that assign ranks
to each literal $\ell$, one defines the rank $r(\omega)$ of a
model $\omega$ as the sum of the ranks of the literals made true
by $\omega$ (i.e., $r(\omega)\doteq\sum_{\omega\vDash\ell}r(\ell)$),
and say that $\omega$ is a best (ranked) model if there is no model
$\omega'$ such that $r(\omega')<r(\omega)$.

Given a theory in d-DNNF, one computes the rank of the best models,
and a best model, in time linear in the size of the d-DNNF.
This computation transforms the DAG of the d-DNNF into a arithmetic
circuit by replacing the AND nodes with `+' and the OR nodes with `min'.
The literal ranking function assigns
values to the leaves of the circuit that are propagated to the
root in linear time. The value of the root is the rank of the
best model \cite{darwiche:weighted}.

In this paper, we exploit the properties of d-DNNFs by constructing
a logical theory whose models encode the valid workflow instantiations
and best models encode the optimal (best) instantiations.
Thus, the combinatorial search is reduced to the computation
of a best model of a logic theory which can be efficiently
performed once the theory is transformed into d-DNNF format.

The paper contains six more sections. The next section illustrates and
motivates the WIP with a simple yet typical example. Section~3 summarizes
the existing related work in three areas of selection of services,
query rewriting, and knowledge compilation.
Then, Sections~4 and 5 describe the architecture of the system and 
report our empirical results over different benchmark problems
respectively.
For the interested reader, Section~6 presents a formal description
of the proposed approach together with an analysis.
The paper concludes with a final discussion in Section~7.

\section{Motivating Example}

Consider a simple flight-information system which contains information
about flights between cities and information about which cities are in
the US. Such a system can be described using LAV with the two abstract
services $\flight(x,y)$ and $\UScity(x)$. The former relates two cities
$x$ and $y$ if there is a direct flight between them, and the latter tells
whether $x$ is a US city or not.

For the concrete services, assume that the available data sources on
the Internet contain the following information:
\begin{enumerate}[--]
\item $\national(x,y)$ relates two US cities that are connected by a direct flight,
\item $\oneway(x,y)$ relates two cities that are connected by a one-way flight,
\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
\item $\flightPA(x)$ tells if there is a direct flight from $x$ to Paris,
\item $\onestopPA(x,y)$ relates $x$ and $y$ if there is a flight from $x$ to Paris
      with a stop at $y$, and
\item $\fromNY(x)$ tells if there is a flight from New York into $x$.
\end{enumerate}
Furthermore, the concrete services are described using the abstract services:
\begin{alignat*}{1}
\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
\end{alignat*}
Now suppose that a user is interested in constructing a workflow able to retrieve
the one-stop round-trip flights from US cities to any city in the world, such that
flights can stop at any city. 
\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
The following conjunctive query represents the workflow that defines this request
in terms of abstract services. The workflow is defined in way that all issues
about the binding of input/output parameters between services had been resolved
such that any instantiation of the abstract services in terms of concrete services
is a valid implementation of the workflow. 
Implementations correspond to compositions of concrete services in which a concrete
service may implement one or more abstract services from the workflow, but each
abstract service can be implemented by exactly one concrete service.
For example, the following composition corresponds to one such implementation.
\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
However, the following two compositions are not valid.
\begin{alignat*}{1}
I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
\end{alignat*}
The first composition is not valid because it maps the workflow variable $y$
into constants \PA\ and \NY\ that denote different cities.
On the other hand, $I''$ does not implement the workflow because the concrete
service $\onestop(x,y)$ does not receive as input, or produce as output, the
middle city where the flight stops and thus is not possible to ensure that 
this city is bound to the city $w$ that is returned by the workflow.

These examples show that a proper workflow instantiations must handle
constants such that not two different constants are mapped to each
other either directly or indirectly via transitivity, and that all
attributes that appear in a join or in the output need to be produced
by the selected concrete services.

The QoS parameters are modeled by annotating the concrete services with
utilities that characterize their behavior which are then aggregated
during instantiation. Thus, as said before, the best instantiation is
the one that minimizes (or maximizes) the aggregation of utilities.

\section{Related Work}

In this section we summarize the existing approaches that provide
solutions to the problems of service selection, query rewriting and
discuss related work in the area of Artificial Intelligence called
knowledge compilation.

\subsection{Service Selection}

The problem of selecting the services that implement an abstract workflow
and best fit the QoS-based criteria is known as the QoS-aware service
selection or composition problem, which has been shown to be NP-hard \cite{Hiroshi2008}.
This problem is a combinatorial optimization problem and several heuristics 
have been proposed to find a relatively good solution in a reasonably short
period of time.

Rahmani et al.\ \cite{rahmani08} present a distance metric-based heuristic
that focus a backward search algorithm; this metric induces an order of the
services in a way that sink nodes are unlikely to be visited.
In a series of papers, Berardi and others \cite{berardi05,berardi08,berardi06}
describe services and workflows in terms of deterministic finite-state machines
that are encoded using Description Logic theories whose models correspond to
solutions of the problem. Although reasoning methods for Description Logics
formalisms could be exploited, scalability or performance of the proposed
solution has not been reported.

Ko et al.\ \cite{myoung08} propose a constraint-based approach that encodes the
non-functional permissible values as a set of constraints whose violation needs
to be minimized; to traverse the space of possibly optimal solutions, a hybrid
algorithm that combines tabu search and simulated annealing meta-heuristics is
implemented. Experimental results show that the proposed solution is able to
scale up to a large number of services and abstract processes.
Cardellini et al.\ \cite{cardellini07} encode one part of the QoS-aware service
composition problem as a Linear Programming problem \cite{cardellini07}.
On the other hand, Wada el at.\ \cite{Hiroshi2008} treat the problem as a 
multi-objective optimization problem where the different QoS parameters are
considered equally important instead of aggregating them into a single function.
Then, a genetic-based algorithm is proposed to identify a set of non-dominated
service compositions that best fit all the QoS requirements.

Alrifai and Risse \cite{alrifaiR09} propose a two-fold solution that uses a
hybrid integer programming algorithm to find the decomposition of global QoS
into local constraints, and then, selects the services that best meet the local
constraints.

Recently, two planning-based approaches have been proposed.
Kuter and Golbeck \cite{kuterG09} extend the SHOP2 planning algorithm to select
the trustworthy composition of services that implement a given OWL-S process model,
while Sohrabi and McIlraith \cite{sohrabiM09} propose a HTN planning-based solution
where user preference metrics and domain regulations are used to guide the planner
into the space of relevant compositions.
Finally, L\'ecu\'e \cite{lecue09} proposes a genetic-based algorithm to identify
the composition of services that best meet the quality criteria for a set of QoS
parameters.

Although these solutions are able to solve the optimization problem and scale up
to a number of abstract processes, none of them are tailored to semantically describe
the services in terms of abstract process, nor to use these descriptions to 
identify the services that implement a given workflow or best meet the user's
non-functional criteria.

\subsection{Query Rewriting}

A number of algorithms have been developed to find the rewritings of a given query;
the most prominent being the bucket algorithm \cite{levy:bucket}, the inverse rules
algorithm \cite{pods:DuschkaG97,Qian96}, the \minicon\ algorithm \cite{pottinger:minicon},
and the \mcdsat\ algorithm \cite{arvelo:aaai06}.
Generally, query rewriting algorithms work in two phases. During the first phase,
the algorithms identify the views that rewrite at least one sub-goal of the query, and
during the second, these partial rewritings are combined into complete rewritings.
The main difference between the algorithms is the criteria used to choose the relevant
views to reduce the space of non-useful rewritings.

The bucket algorithm reduces the number of possibilities by just considering each
sub-goal in the query in isolation, and selecting the views that are able to produce
at least the attributes projected by the query. Since the attributes involved in the
joins in the query are not verified, a large number of rewritings comprised of Cartesian
products may be generated.

The Inverse Rules algorithm constructs a set of rules that invert the view definitions
and establish how to compute tuples for the database relations from the tuples of the
views. Similarly to the bucket algorithm, it can produce a large number of non-useful
rewritings.

The \minicon\ algorithm overcomes the limitations of the previous algorithms by identifying
only views that rewrite a set of the sub-goals of the query, and that can be combined with
the rest of the sub-goals. The key idea is to identify the mappings between the variables in
each sub-goal to the variables in one or more sub-goals in the views, in a way
that, join variables in the query are mapped to join variables in the body of a view or to
the distinguished variables of the view. Mappings between variables and sub-goals are
represented in the so-called MiniCon Descriptions (MCDs) \cite{pottinger:minicon}.

Finally, the \mcdsat\ algorithm is able to identify the query rewritings of a query by
translating the problem of rewriting into the problem of enumerating the models of a
propositional theory whose models are in correspondence with the rewritings of the query.
The algorithm exploits the properties of d-DNNFs to efficiently enumerate the models of
the theory.
The \mcdsat\ algorithm has demonstrated to scale better than the \minicon\ algorithm over
a large number of benchmarks often showing performance improvements of several orders of
magnitude. However, the \mcdsat\ algorithm was not designed for rewriting problems
involving explicit constants, nor to compute the best rewritings with respect to a given
utility function or cost model. In this paper, we propose an extended encoding that 
overcomes these limitations and apply the encoding to the Workflow Instantiation Problem.

\subsection{Knowledge Compilation}

Knowledge compilation is the area in AI concerned with the problem of mapping
logical theories into suitable fragments that make certain desired operations
tractable \cite{cadoli:compilation}. Different compilation languages have been
defined, for instance, Ordered Binary Decision Diagrams (OBDDs) \cite{bryant:obdd},
Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation
Normal Form (DNNF) \cite{darwiche:map}. In this work, we make use of the properties
of the deterministic DNNFs (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable
and efficient solution to the service selection problem.

\begin{figure}[t]
\centering
\includegraphics[width=.7\textwidth]{odd}
\caption{A decomposable and deterministic NNF.}
\label{fig:dnnf}
\end{figure}

A Negation Normal Form (NNF) theory is constructed from literals using only conjunctions
and disjunctions \cite{barwise:handbook}, and it can be represented as a directed
acyclic graph in which the leaves are labeled with literals and the internal nodes
are labeled with $\land$ and $\lor$; see Fig.~\ref{fig:dnnf} for an example.
This fragment is universal meaning that for every logical formula there is an
equivalent one in NNF format.
An NNF is said to be decomposable (DNNF) \cite{darwiche:d-dnnfs} if for each conjunction
$\bigwedge_i\phi_i$, the set of variables in each conjunct are pairwise disjoint; i.e,.
$Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i\neq j$.
A DNNF supports a number of operations in polytime in the size of its DAG.
For example, we can test whether a DNNF is satisfiable by a single bottom-up pass
over its DAG in linear time.
A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each
disjunction $\bigvee_i\phi_i$, the disjuncts are pairwise logically contradictory;
i.e., $\phi_i\land\phi_j\equiv\textbf{false}$ for $i\neq j$.
The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable and deterministic.
A d-DNNF supports model counting in polytime in the size of its DAG, and model
enumeration in polytime in the size of the output.
Furthermore, given a literal ranking function $r$, one can compute the rank
of the best model in polytime for DNNFs \cite{darwiche:weighted}.

The fragments DNNF and d-DNNF are universal yet translating a CNF theory 
to DNNF format has an exponential cost in the worst case. This translation
is referred to as compilation in the field. There is a publicly available
compiler, called c2d,\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}
that perform this compilation process and that make
use of modern SAT techniques such as conflict-directed backtracking,
clause learning and caching \cite{darwiche:compiler}.
This compiler incurs in the worst case in exponential space in a parameter
called the width of the decomposition tree that is related to the ``connectivity''
of the CNF theory. However, in our experiments, the CNF theories that are
compiled are of low width.

%DNNFs had also been used elsewhere to tackle other combinatorial optimization
%problems such as problems in planning \cite{palacios:icaps05} and diagnosis
%\cite{REF}.

\begin{figure}[t]
\centering
\includegraphics[width=.9\textwidth]{architecture}
\caption{System Architecture}
\label{fig:architecture}
\end{figure}

\section{System Architecture}

We use an architecture that is comprised of a Catalog of service descriptions,
the Encoder, the compiler c2d, the best model Finder, and the Decoder.
Figure~\ref{fig:architecture} depicts the overall architecture of the system.
In this framework, an instance of Workflow Instantiation Problem is defined as an
abstract workflow represented by a conjunctive query on abstract services which
is given as input together with a set set of concrete services defined by views
of abstract services.

@@

An input instance of WIP is encoded as a CNF theory whose models corresponds to
the instantiations of the workflow by the Encoder. The compiler c2d, an off-the-shelf
component, compiles the CNF formula into d-DNNF.
The Encoder translates the WIP instances into CNF theories, that are then converted
into d-DNNF using c2d. The Finder computes a best model given the QoS parameters
in linear time in the size of the resulting d-DNNF. It is important to remark,
that the compilation process needs to be performed only once as it does not
depend on the value of the QoS parameters. Thus, even if the compilation happens
to be costly in terms of time, this cost can be amortized since the resulting
d-DNNF can be used to find best instantiations with respect to multiple values
of the QoS parameters.
Finally, the Decoder translates the best model returned by the Finder into
a workflow instantiation that solves the WIP.

Given a CNF that encodes a WIP, its d-DNNF is a compact representation of
all the workflow instantiations. That is, one can generate in a backtrack-free
manner all the instantiations of the workflow. If the user is interested in
a best instantiation given the QoS parameters, then it can be computed in
linear time in the size of the d-DNNF. If the user is interested in the 
all the best instantiations, these can be computed in linear time in the
number of them. Finally, if the user is interested in all instantiations,
these can be also computed in linear time in the number of them.
In the latter two cases, if such number is exponential (in the size of the
input), the enumeration of the instantiations is also exponential but
this complexity is intrinsic to the problem and thus cannot be avoided.

In order to make our results more accessible to a general audience, we decided
to present the experimental results in the next section, and leave the formal
and theoretical results for the following section. In this way, the interested
reader may skip the formal details of the approach on a first reading.

\section{Experimental Results}

We conducted an empirical analysis on three benchmarks.
All the experiments were run on a desktop machine with an Intel
Core 2 Duo 2GHz CPU and 4Gb of memory, and the time was measured
with the Unix time command.

The objective of the experiment is to assess the performance
of the proposal on varying conditions. The main benefit of our approach
is that one can compile the logical theory for a problem instance
and then calculate all the instantiations, or the best ones, any 
number of times. The cost model for finding best instantiations
can be changed with no need to recompile the logical theory.
Therefore, the time complexity of our approach is basically the
time to encode the WIP as a CNF plus the time to compile the
CNF into a d-DNNF and the time to decode the models. 
The times to encode and decode are negligible compared to
the time to compile the CNF. Because of this, we focus on the
time to compile the benchmark problems.

The first benchmark consists of problems for air-travel queries.
Concrete services are of the form $V_i(x,y)\qrule\ \flight(x,y,\AL_i)$
where $\AL_i$ is a constant that denotes the name of an airline, 
and $\flight(x,y,\AL_i)$ relates the cities $x$ and $y$ such that
there is a flight between $x$ and $y$ served by $\AL_i$. This concrete
service is assumed to return all flights between two cities with an
specific airline. The workflow has the form 
\[ W(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,z),\,\flight(x_1,x_2,z),\,\ldots,\,\flight(x_n,\NY,z)\,. \]

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{plots/plot1}
\caption{Compilation times for experiments I and II for different
number of goals and different number of views. The plots are in
logarithmic scale, and the time is in seconds.}
\label{fig:plot1}
\end{figure}

The benchmark includes instances for workflows with 2 to 5 sub-goals and
sets of 10 to 100 concrete services. The results for the compilation are
shown in panel (a) of Fig.~\ref{fig:plot1}. This is a plot in logarithmic
scale that suggests a sub-exponential behavior. In any case, the results
show good performance since realistic instances of the problem (sets of 100
airlines with 5-stop flights) can be compiled in 328 seconds.
The size in disk of the d-DNNF for 100 airlines and 5-stop flights is 3.4Mb.
On this d-DNNF, the best model can be computed in 0.29 seconds,
and the enumeration of all models in 0.47 seconds.

In an attempt to induce an exponential growth in the compilation time,
in the second benchmark we add a second concrete service for each airline.
This modification increases the number of valid instantiations from linear
to exponential since each leg of a flight can now be instantiated by two
concrete services and thus a flight with $n$ legs may have up to $2^n$
instantiations.
We ran the compiler for instances comprising the same number of workflow
sub-goals and total number of concrete services. The results plotted in
logarithmic scale are shown in panel (b) of Fig.~\ref{fig:plot1}.

These tests show good performance for this type of problems, but they do not
involve concrete services with multiple sub-goals. We therefore designed a
third experiment that consists unstructured, randomly generated instances.
Each instance contains three variables per abstract service, ten distinct
variables and ten distinct constants, six sub-goals in the workflows,
2 to 5 sub-goals in the concrete services, and a varying number of services.
The chance that an argument of an abstract service is bound to a constant is 50\%.
The results are depicted in Fig.~\ref{fig:plot3}.
The compilation time for these instances does not grow monotonically
since they are randomly generated. The same happens for the size of
the theories and the number of models. For example, the d-DNNF for 
a problem with 45 views each with 5 sub-goals is of size 5.1Mb and has
$1.26\times 10^8$ models. The time to find the best model for
this d-DNNF is 0.46 seconds while the time to enumerate all models
is about 17 hours.

\begin{figure}[t]
\centering
\includegraphics[width=.8\textwidth]{plots/plot3}
\caption{Compilation times for experiments III for different number of goals
and different number of views. The plots are in logarithmic scale, and the time
is in seconds.}
\label{fig:plot3}
\end{figure}

These are preliminary experiments, yet the results show that the
proposed approach efficiently scales for problems with several goals
and views. We believe that these results are encouraging and motivate
us to continue this research. We plan to conduct additional experiments
with other types and sizes of workflows and, when possible, compare with
other approaches.

\section{Formalization}

>>>>

 \documentclass{llncs}
 \usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{url}

@@

\begin{document}
\allowdisplaybreaks
\title{SatWins: A Sat-based tool to Optimal Instantiations of Abstract Workflows}
\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
\institute{Departamento de Computaci\'on \\
           Universidad Sim\'on Bol\'{\i}var \\
           Caracas 89000, Venezuela \\
           \url{{idaniel,mvidal,bonet}@ldc.usb.ve}}
\maketitle

\begin{abstract}
We present SatWins a system to efficiently solve the abstract workflow
instantiation problem that is firmly grounded on logic.
SatWins adopts the Local-As-View approach in which
the functionality of concrete services is described using views
of abstract services, the quality of an instantiation as an
overall utility function that combines the different QoS parameters,
and the abstract workflows as conjunctive queries on abstract services.
We discuss the encoding of the workflow instantiation problem as a logical theory whose models
are in correspondence with the instantiations of the workflow,
and the best ranked models are in correspondence with the
optimal instantiations of the workflow. We demonstrate how by exploiting known properties of logical theories in 
d-DNNF format, SatWins provides an efficient and scalable solution
to the workflow instantiation problem. 
\end{abstract}

\section{Introduction}
Emerging technologies support the basis for the wealth of available Web resources, as well as current users' tendency to rely more on automatic methods for handling existing Web resources or services composed in complex workflows. The execution of an abstract workflow involves the instantiation
of the abstract services into concrete services that meet the workflow
functional and non-functional requirements. This instantiation
process can be seen as a search for a target instantiation on
the combinatorial space of all valid instantiations.
Thus, one is interested in efficient techniques for performing
this search that are able to scale up as the number of concrete
services or the complexity of the workflow increases. We call the problem of instantiating a given abstract workflow with concrete services, from a given pool of concrete services,
such that certain QoS demands are meet, the Workflow Instantiation
Problem (WIP).

In this paper, we describe SatWins a system that considers
 a restricted version of WIP that adopts
the Local-As-View (LAV) approach \cite{levy:bucket} to specify the semantics of the concrete services as views of abstract services; this representation is similar to the
one that is semi-automatically generated for the DEIMOS system
\cite{AmbiteISWC09}. Additionally,  abstract workflows correspond  to conjunctive
queries on the abstract services, and  the quality
of an instantiation is an overall utility function that combine the
different QoS parameters of the selected concrete services.
In this version of WIP, the rules that define workflows (conjunctive queries)
and concrete services (views) are created in a way that all the 
functional restrictions on pre- and post-conditions of services
and their combinations are satisfied, and the QoS measures are
represented by annotating each concrete service description with
a real number that represent the overall QoS utility of the service.

In addition to provide a novel approach to describe the WIP problem, SATWins is able to  efficiently identify an abstract workflow instantiations. 
Given a WIP, a logical theory is constructed such that each model of
the theory encodes a valid instantiation, and thus all valid instantiations
are obtained by enumerating all models of the theory. This enumeration
can be efficiently performed if the logical theory is in a certain 
normal form called deterministic and decomposable negation normal
form (d-DNNF) \cite{darwiche:d-dnnfs}. Thus, the approach consists in
transforming (called compiling in the field of knowledge compilation)
the logical theory into d-DNNF format for enumerating its models efficiently; also, the  enumeration 
the best ranked models can also be performed efficiently on d-DNNF and the complexity is independent of the utility function used to rank the optimal instantiation.

In this demonstration, we  attempt to show the following key aspects of the approach implemented by SATWins. We will use a real-world example based on a simple flight-information system.
\begin{itemize}
\item We show the expressiveness of representing concrete services as views on abstract services.
\item We demonstrate the performance of the proposed techniques by computing the optimal instantiation of a given WIP in terms of different utility functions.  
\item We show the scalability of our solution by demonstrating how SATWins is able to enumerate e realistic instances of the problem in less of one second.
\end{itemize}

\section{SatWins Architecture}
 The SatWins architecture  is comprised of a Catalog of service descriptions,
the Encoder, the compiler c2d, the best model Finder, and the Decoder.
Figure~\ref{fig:architecture} depicts the overall architecture of the system.
In this framework, an instance of Workflow Instantiation Problem is defined as an
abstract workflow represented by a conjunctive query on abstract services which
is given as input together with a set set of concrete services defined by views
of abstract services.

\begin{figure}[t]
\centering
\includegraphics[width=.9\textwidth]{architecture}
\caption{The SatWins Architecture}
\label{fig:architecture}
\end{figure}

@@

An input instance of WIP is encoded as a CNF theory whose models correspond to
the instantiations of the workflow by the Encoder. 
The compiler c2d, an off-the-shelf
component, compiles the CNF formula into d-DNNF.
The Encoder translates the WIP instances into CNF theories, that are then converted
into d-DNNF using c2d. The Finder computes a best model given the QoS parameters
in linear time in the size of the resulting d-DNNF. It is important to remark
that the compilation process needs to be performed only once as it does not
depend on the value of the QoS parameters. Thus, even if the compilation happens
to be costly in terms of time, this cost can be amortized since the resulting
d-DNNF can be used to find best instantiations with respect to multiple values
of the QoS parameters.
Finally, the Decoder translates the best model returned by the Finder into
a workflow instantiation that solves the WIP.

Given a CNF that encodes a WIP, its d-DNNF is a compact representation of
all the workflow instantiations. That is, one can generate in a backtrack-free
manner all the instantiations of the workflow. If the user is interested in
a best instantiation given the QoS parameters, then it can be computed in
linear time in the size of the d-DNNF. If the user is interested in the 
all the best instantiations, these can be computed in linear time in the
number of them. Finally, if the user is interested in all instantiations,
these can be also computed in linear time in the number of them.
In the latter two cases, if such number is exponential (in the size of the
input), the enumeration of the instantiations is also exponential but
this complexity is intrinsic to the problem and thus cannot be avoided. Figure \~ref{fig:snapshot} shows a snapshot of an input instance of WIP and the functionality of SatWins.

\begin{figure}[t]
\centering
\includegraphics[height=30mm,width=.7\textwidth]{demo}
\caption{A snapshot of the SatWins interface}
\label{fig:snapshot}
\end{figure}

\section{Demonstration Use Cases}
In this demonstration, we will go through the different steps of the proposed approach. We illustrate our approach on  a simple flight-information system which contains information
about flights between cities and information about which cities are in
the US. Such a system can be described using LAV with the two abstract
services $\flight(x,y)$ and $\UScity(x)$. The former relates two cities
$x$ and $y$ if there is a direct flight between them, and the latter tells
whether $x$ is a US city or not. We also consider different sets of concrete services; functionality, complexity and size will be changed during the demonstration to show expressiveness, performance and scalability of our approach. An example of some of the concrete services to be used is as follows:
\begin{enumerate}[--]
\item $\national(x,y)$ relates two US cities that are connected by a direct flight,
\item $\oneway(x,y)$ relates two cities that are connected by a one-way flight,
\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
\item $\flightPA(x)$ tells if there is a direct flight from $x$ to Paris,
\item $\onestopPA(x,y)$ relates $x$ and $y$ if there is a flight from $x$ to Paris
      with a stop at $y$, and
\item $\fromNY(x)$ tells if there is a flight from New York into $x$.
\end{enumerate}
The descriptions of these concrete services using the abstract services:
\begin{alignat*}{1}
\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
\end{alignat*}

In addition, the following abstract workflow to retrieve
the one-stop round-trip flights from US cities to any city in the world, such that
flights can stop at any city, is represented as a conjunctive query on the abstract services as follows: 
\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]

Any instantiation of the abstract services in terms of
concrete services is a valid implementation of the workflow or concrete service compositions.
For example, the following composition corresponds to one such implementation.
\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
However, the following composition is not valid.
\begin{alignat*}{1}
I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
\end{alignat*}

Using this domain, we will demonstrate the following scenarios:
\begin{itemize}
\item To show the scalability of the proposed approach, we use the behavior of SATWins in three different benchmarks. In all the cases, we show time for compilation, size of the  d-DNNF, and time to enumerate all the models as well as the optimal one. We start with a benchmark of instances for workflows with 2 to 5 sub-goals and sets of 10 to 100 concrete services. Although, the observed time suggests a sub-exponential behavior,  in any case, the results
show good performance. For instance, sets of 100
airlines with 5-stop flights can be compiled in 328 seconds; the size in disk of the d-DNNF is 3.4Mb;
the best model can be computed in 0.29 seconds, and the enumeration of all models in 0.47 seconds.
Then, in an attempt to induce an exponential growth in the compilation time, we add a second concrete service for each airline and we will show that SATWins behavior remains similar. Finally, we  involve concrete services with multiple sub-goals which are randomly generated.  The compilation time for these instances does not grow monotonically
since they are randomly generated. The same happens for the size of
the theories and the number of models. For example, the d-DNNF for 
a problem with 45 views each with 5 sub-goals is of size 5.1Mb and has
$1.26\times 10^8$ models. The time to find the best model for
this d-DNNF is 0.46 seconds while the time to enumerate all models
is about 17 hours.
\item To show the performance of the approach and reusability of existing d-DNNF, we will allow users to choose different utility metrics to compute the best solution.  We will show that the existing d-DNNF can be used to compute the best model with different utility functions, while the execution time remains the same for the different functions. 
\end{itemize}

\section{Conclusions}
In this demonstration, we will present a Local-As-View (LAV)-based approach \cite{levy:bucket} to WIP. We will show how by using simple conjunctive rules the functionality of concrete services and abstract workflows can be defined. In addition, we will show how a propositional-based approach can be used to represent instances of WIP and different scenarios where the performance and scalability of the technique will be observed by the users.  We will demonstrate that the approach can be applied to
real-sized workflow problems, while the whole approach is only possible
when the compilation of the CNF theory into d-DNNF format succeeds.
\bibliographystyle{abbrv}
\bibliography{ref}
 \end{document}
